---
title: " Who is Canberra - Generate Ancestry Data"
output: html_document
---

### Use
This is an R markdown document. Executing 'knit' on the .Rmd version of this document will generate an HTML document containing this text and the R code, and execute the code generating files in the format expected by the Who is Canberra application.

If you want to Knit the script you need R and the knitr package (and its' dependencies), the "rjson" package, and an active internet connection to download the data from stat.abs.gov.au

```{r}
require("rjson")
require(RCurl)
```

### Download ABS data

The following URL calls the ABS ITT query interface for the Census 2011 Ancestry by country of Parent's Birth dataset (ABS_CENSUS2011_B08), selecting only the regions of interest, and the total responses for each ancestry.

Regions selected:
  SA Level 4: 
   - ACT 801
  SA Level 3: 
  - Belconnen 80101, 
  - Gungahlin 80104
  - North Canberra 80105
  - South Canberra 80106
  - Tuggeranong 80107
  - Weston Creek 80108
  - Woden 80109
  - Queenbeyan 10102

```{r}
dataURL <- "http://stat.abs.gov.au/itt/query.jsp?method=GetGenericData&datasetid=ABS_CENSUS2011_B08&and=MEASURE.TOT,ANCP.&or=REGION.10102,REGION.801,REGION.80101,REGION.80104,REGION.80105,REGION.80106,REGION.80107,REGION.80108,REGION.80109&format=json"

``` 

Read the data from the URL and do a dance to extract the data from the Json provided by the API

```{r}

fullJson <- fromJSON(getURL(dataURL))
smallJson <- fullJson$series

tempmatrix <- matrix(data = NA, nrow = 306, ncol = 3)
for (i in 1:306){
tempmatrix[i,] <- rbind(smallJson[[i]]$concepts[[2]][[2]],smallJson[[i]]$concepts[[5]][[2]],smallJson[[i]]$observations[[1]][[2]])
}

fullFrame <- data.frame(tempmatrix)
colnames(fullFrame)<-c("ANCP","REGION","Value")

```

The ancestry codes we aren't interested in are subsetted out using ABS ANCP codes:
Other(d) OTH
Total responses(c) TOT
Ancestry not stated Z

```{r}
workingFrame <- fullFrame[!(fullFrame$ANCP == "OTH") & !(fullFrame$ANCP == "TOT")  & !(fullFrame$ANCP == "Z"),]

transform(workingFrame, Value = as.numeric(Value))

```

Turning the ANCP codes into countries makes the output useful, albeit this is a bit kludgy at the moment.

```{r}

## This line is the quickest way to sort out a peculiarity of how R treats strings and factors in dataframes in this case
levels(workingFrame$ANCP) <- c(levels(workingFrame$ANCP), "Australian","Australian Aboriginal","Chinese","Croatian","Dutch","English","Filipino","French","German","Greek","Hungarian","Indian","Irish","Italian","Korean","Lebanese","Macedonian","Maltese","Maori","New Zealander","Polish","Russian","Scottish","Serbian","Sinhalese","South African","Spanish","Turkish","Vietnamese","Welsh","Total persons(c)")


for (i in 1:nrow(workingFrame)){

  if (workingFrame[i,"ANCP"] == "1101") {
    workingFrame[i,"ANCP"] <- "Australian"
  }
  if (workingFrame[i,"ANCP"] == "1102") {
    workingFrame[i,"ANCP"] <- "Australian Aboriginal"
  }
    if (workingFrame[i,"ANCP"] == "6101") {
    workingFrame[i,"ANCP"] <- "Chinese"
  }
    if (workingFrame[i,"ANCP"] == "3204") {
    workingFrame[i,"ANCP"] <- "Croatian"
  }
    if (workingFrame[i,"ANCP"] == "2303") {
    workingFrame[i,"ANCP"] <- "Dutch"
  }
    if (workingFrame[i,"ANCP"] == "2101") {
    workingFrame[i,"ANCP"] <- "English"
  }
  if (workingFrame[i,"ANCP"] == "5201") {
    workingFrame[i,"ANCP"] <- "Filipino"
  }
  if (workingFrame[i,"ANCP"] == "2305") {
    workingFrame[i,"ANCP"] <- "French"
  }
  if (workingFrame[i,"ANCP"] == "2306") {
    workingFrame[i,"ANCP"] <- "German"
  }
  if (workingFrame[i,"ANCP"] == "3205") {
    workingFrame[i,"ANCP"] <- "Greek"
  }
  if (workingFrame[i,"ANCP"] == "3304") {
    workingFrame[i,"ANCP"] <- "Hungarian"
  }
  if (workingFrame[i,"ANCP"] == "7106") {
    workingFrame[i,"ANCP"] <- "Indian"
  }
    if (workingFrame[i,"ANCP"] == "2201") {
    workingFrame[i,"ANCP"] <- "Irish"
  }
    if (workingFrame[i,"ANCP"] == "3103") {
    workingFrame[i,"ANCP"] <- "Italian"
  }
    if (workingFrame[i,"ANCP"] == "6902") {
    workingFrame[i,"ANCP"] <- "Korean"
  }
    if (workingFrame[i,"ANCP"] == "4106") {
    workingFrame[i,"ANCP"] <- "Lebanese"
  }
    if (workingFrame[i,"ANCP"] == "3206") {
    workingFrame[i,"ANCP"] <- "Macedonian"
  }
    if (workingFrame[i,"ANCP"] == "3104") {
    workingFrame[i,"ANCP"] <- "Maltese"
  }
    if (workingFrame[i,"ANCP"] == "1201") {
    workingFrame[i,"ANCP"] <- "Maori"
  }
    if (workingFrame[i,"ANCP"] == "1202") {
    workingFrame[i,"ANCP"] <- "New Zealander"
  }
    if (workingFrame[i,"ANCP"] == "3307") {
    workingFrame[i,"ANCP"] <- "Polish"
  }
      if (workingFrame[i,"ANCP"] == "3308") {
    workingFrame[i,"ANCP"] <- "Russian"
  }
      if (workingFrame[i,"ANCP"] == "2102") {
    workingFrame[i,"ANCP"] <- "Scottish"
  }
      if (workingFrame[i,"ANCP"] == "3213") {
    workingFrame[i,"ANCP"] <- "Serbian"
  }
      if (workingFrame[i,"ANCP"] == "7115") {
    workingFrame[i,"ANCP"] <- "Sinhalese"
  }
      if (workingFrame[i,"ANCP"] == "9215") {
    workingFrame[i,"ANCP"] <- "South African"
  }
        if (workingFrame[i,"ANCP"] == "3106") {
    workingFrame[i,"ANCP"] <- "Spanish"
  }
        if (workingFrame[i,"ANCP"] == "4907") {
    workingFrame[i,"ANCP"] <- "Turkish"
  }
        if (workingFrame[i,"ANCP"] == "5107") {
    workingFrame[i,"ANCP"] <- "Vietnamese"
  }
        if (workingFrame[i,"ANCP"] == "2103") {
    workingFrame[i,"ANCP"] <- "Welsh"
  }
        if (workingFrame[i,"ANCP"] == "TOTP") {
    workingFrame[i,"ANCP"] <- "Total persons(c)"
  }
}
```
  
### Subsetting the data by region

The engine expects data as a single .csv file for each region, with no headers and with each row a country followed by the value. The file should also have a final row providing the total by which the other rows are divided to indicate the ratio.

This code extracts the rows and columns needed for the ACT data. The ABS Region codes provided above are used to subset the appropriate data.

```{r}
ACTFrame <- workingFrame[workingFrame$REGION == 801,c("ANCP","Value")]
```

This code extracts the data for the Statistical Area Level 3 catagories desired: Belconnen, Gungahlin, North Canberra, South Canberra, Tuggeranong, Weston Creek, Woden and Queenbeyan

```{r}
belFrame <- workingFrame[workingFrame$REGION == 80101,c("ANCP","Value")]
gungFrame <- workingFrame[workingFrame$REGION == 80104,c("ANCP","Value")]
northFrame <- workingFrame[workingFrame$REGION == 80105,c("ANCP","Value")]
southFrame <- workingFrame[workingFrame$REGION == 80106,c("ANCP","Value")]
tuggFrame <- workingFrame[workingFrame$REGION == 80107,c("ANCP","Value")]
westonFrame <- workingFrame[workingFrame$REGION == 80108,c("ANCP","Value")]
wodenFrame <- workingFrame[workingFrame$REGION == 80109,c("ANCP","Value")]
queenFrame <- workingFrame[workingFrame$REGION == 10102,c("ANCP","Value")]
```

### Writing output files
The data frames are written to .csv files named in the format "datatype (plural of region).csv" to conform to engine expectations.

```{r}
## write data frames
write.table(ACTFrame, file = "./Ancestry (Canberrans).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(belFrame, file = "./Ancestry (Belconnenites).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(gungFrame, file = "./Ancestry (people in Gungahlin).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(northFrame, file = "./Ancestry (North Canberrans).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(southFrame, file = "./Ancestry (South Canberrans).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(tuggFrame, file = "./Ancestry (people in Tuggeranong).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(westonFrame, file = "./Ancestry (people in Weston Creek).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(wodenFrame, file = "./Ancestry (Wodenites).csv", row.names = FALSE, sep = ",", col.names = FALSE)
write.table(queenFrame, file = "./Ancestry (Queenbeyanites).csv", row.names = FALSE, sep = ",", col.names = FALSE)
```

The data files are now ready to load into the Who is Canberra application.
